### 2014-04-21 patchy-selection-paper.tex
draft! cleaned up!

Let $p(x)$ be the probability that a new mutant $B$ allele arising at location $x$ 
relative to the patch fixes within the poor habitat patch.
As shown in Appendix \ref{apx:elliptic_integrals} XXX,
the function $p(x)$ can be found (in one dimension) in terms of Jacobi elliptic functions,
but the precise expressions are not particularly helpful.
The total successful mutational influx per generation is then given by the integral of $p(x)$ over the entire species range,
% $\int \rho \mu p(x) dx$,
and hence depends in a complicated way on the relationship of patch width and selection coefficients,
but still scales linearly with the mutational influx density $\rho \mu$.
If the width of the patch is large relative to $\sigma/\sqrt{2s_m}$, 
then a reasonable approximation is to assume that $p(x) = p_e$ within the patch, and $p(x) = 0$ otherwise,
for a patch of area $A$ approximating $\int p(x) dx \approx 2 s_p A / \xi^2$. 
This holds for large patches, in which the rate at which mutations arise within the patch is larger 
than the rate at which they appear nearby;
we examine this more generally via simulation in Appendix \ref{apx:establishment_sims}.

@@

Migration--selection balance ensures that 
there will be some number of $B$ alleles present in the regions where they are disadvantageous,
but only very few, and rarely, far away from the patch where $B$ is established.
Denote the expected frequency of allele $B$ at displacement $x$ relative to the patch by $q(x)$.
Following \citet{slatkin1973geneflow}, one can write down a differential equation to which $q(x)$ is the solution,
which in one dimension can be solved to give $q(x)$ in terms of Jacobi elliptic functions (see Appendix \ref{apx:elliptic_integrals}).
More general, asymptotics of the solution implies that for large $|x|$ the mean occupation frequency decays exponentially, with a square-root correction in two dimensions (see Appendix \ref{apx:asymptotics}): 
\begin{align} \label{eqn:eqfreq}
    q(x) \approx C |x|^{-(d-1)/2} \exp( - |x| \sqrt{2 s_m} / \sigma) \qquad \text{for large $|x|$},
\end{align}
where $d$ is the dimension ($d=1$ or $2$), and $C$ and $r_0$ are constants depending on the geographic shape of the populations and the selection coefficients
(but note that $r_0$ is only necessary in $d=2$).
In applications we fit $C$ and $r_0$ to the data;
to get concrete numbers from theory we take $C=1$ and $r_0=0$ if necessary.
From simulations, $C$ seems to be not too far from 1 in practice (see Appendix \ref{apx:patchy_sims}).

This expected frequency $q(r)$ is the \emph{time-averaged} occupation frequency,
or in other words, the total number of $B$ alleles found at distance $r$ across $T$ generations, divided by $T$.
Therefore, if we look at a small area of size $\epsilon$, 
then $q(r) \epsilon$ is equal to the probability of finding any $B$ alleles there,
multiplied by the mean number of such, if there are any.
This points to an important fact: 
close to the original patch, the ``equilibrium frequency'' $q(r)$ describes well the density of the $B$ allele at most times,
but far away from the patch, 
the equilibrium is composed of rare excursions of a family of $B$ alleles, interspersed by long periods of absence.
The relevant time scale on which $B$ alleles migrate between patches is given by the rate of appearance of such families.
To compute the rate of establishment by such migration,
we use properties of subcritical branching processes and Brownian motion,
combined with equation \eqref{eqn:eqfreq},
to obtain an approximation that is valid if the distance $r$ is large (at least five times $\sigma/\sqrt{s_m}$, say).

@@

Now we turn to the expression \eqref{eqn:gestalt_migrate} for the outflux of successful migrants.
Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch,
which we assume is circular, has relatively small area $A$ and is located near position $y$.
% The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-p_e(t))$,
% multiplied by the probability that the family reaches the patch, given survival.
% If $t$ is large then as in the previous section, $(1-p_e(t)) \approx e^{-\alpha t}/\E[K]$.
It would be possible to analyze directly the probability that a given migrant family establishes in $A$,
as in \citet{barton1987probability,branchingbrownian};
but we make a simpler approximation, approximating this by the chance that the ``trunk'' hits the new patch,
multiplied by the chance that at least member of the family escapes demographic stochasticity to establish.
Write $h(x,y)$ for the probability that the Brownian trunk
hits the new patch centered at $y$ before the family dies out,
and $f(A)$ for the chance that the family manages to establish in the new patch,
given that it successfully arrives.
As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is an integral of $h(x,y)$:
\begin{align}\label{eqn:migrate_integral}
  \migrate(y) &= \rho \int_{\partial B(r_0)} \gamma(x) h(x,y) f(A) dx .
\end{align}

Intuitively, $h(x,y)$ counts each family once if it hits the patch,
while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
Since the expected total time spent in the patch does not depend on the distance at which the family started
if we know that it survived to hit the patch,
it is intuitively reasonable that $u(x,y)$ and $h(x,y)$ are simply related.
Using this fact,
and that $(1-p_e(t)) \approx e^{-\alpha t}/\E[K]$,
it turns out (see Appendix~\ref{apx:hitting_occupation}) that
\begin{align} 
  h(x,y) = \frac{ s(A) }{ \E[K] } u(x,y),
\end{align}
where $s(A)$ is the total expected occupation time of the patch given the family arrives,
and recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a constant factor
(at least asymptotically for large distances),
we now know that the outflux of successful migrant families is equal to the mean occupation frequency
multiplied by $s(A) f(A) / \E[K]$.
The probability $f(A)$ that the family does not die out once it arrives at the new patch
is approximately $\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$.
% If we let $w$ be the width of the patch, then the remaining term is
% \begin{align}
%   s(w) \simeq \frac{1}{2\alpha} \left( 1 - \left( \frac{\pi w \sqrt{2 \alpha}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2\alpha}/\sigma} \right).
% \end{align}
Combining these terms with the expression for $s(A)$ from Appendix~\ref{apx:hitting_occupation},
\begin{align}
  \label{eqn:migrate}
  \migrate(y) \approx q(y) \times \rho \frac{ s_p  }{ \alpha \xi^2 } 
  % \left( 1 - \left( \frac{\pi w \sqrt{2 \alpha}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2\alpha}/\sigma} \right).
\end{align}
For patches that are small relative to $\sigma/\sqrt{2s_m}$, there is an additional correction term (see Appendix~\ref{apx:hitting_occupation}),
but at this point we would be violating various other assumptions above, so we omit it.

%%%%%%% %%%%%%%%%%%
\subsubsection*{Mean occupation density}

XXX OMIT FROM HERE? XXX

First, we compute the second term in equation \eqref{eqn:gestalt_q}, (mean occupation density of a family at $x$),
by which we mean the expected total number of offspring of a single family that ever live at distance $r$ from the patch.
In two dimensions, this depends on where the family began, 
and right-hand side of equation \eqref{eqn:gestalt_q} should be integrated over this starting point.
Here we'll write $x_0$ for the starting point, and $u(x_0,x)$ for the mean occupation density at $x$.
Concretely, suppose that $Z$ is a spatial branching process as above, started with a single individual at $x_0$,
defined now so that $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the set $A$.
The mean occupation density of $Z$ is then defined to be the density of $\E[\int_0^t Z_t dt]$,
i.e.\ the smooth function $u(x_0,x)$ that, integrated over a set $A$, 
gives the expected total number of individuals ever living in $A$:
\begin{align} \label{eqn:occupation_density_defn}
    \E\left[ \int_0^\infty Z_t(A) dt \right] = \int_A u(x_0,x) dx .
\end{align}
Since a family has on average $(1-s_m)^t$ members in generation $t$,
and the marginal distribution of any one surviving individual is approximately Gaussian with variance $t \sigma^2$,
exchanging the expectation and the integral shows that
the density $u(x_0,x)$ is equal to the Green's function for Brownian motion killed at rate $1-s_m$,
\begin{align} \label{eqn:occupation_density_solution}
    u(x_0,x) = \int_0^\infty (1-s_m)^t p_{\sigma^2 t}(x_0,x) dt
\end{align}
where $p_t(x,y) = (2 \pi t)^{-d/2} \exp(-\|x-y\|^2/2t)$ is the transition density for standard Brownian motion.
The integral can be expressed in terms of modified Bessel functions:
using \citet{gradshteyn2007table} (3.471.9 and 8.451.6), the mean occupation density is,
with $\alpha = -\log(1-s_m)$ and $d$ the dimension,
\begin{align}
u(x_0,x) &= \frac{\alpha^{(d-2)/2}}{\sigma^d \pi^{d/2}}\left(\frac{ \sqrt{2 \alpha} \|x-x_0\|}{\sigma} \right)^{(2-d)/2} K_{(2-d)/2}\left( \sqrt{2\alpha} \frac{\|x-x_0\|}{\sigma} \right) \\
         &= 
    \exp\left(- \sqrt{2\alpha} \frac{\|x-x_0\|}{\sigma} \right)
    \left\{
    \frac{\alpha^{(d-2)/2}}{\sigma^d 2 \pi^{(d-1)/2}} \left( \frac{\sqrt{2 \alpha} \|x-x_0\|}{\sigma} \right)^{(1-d)/2} 
+ O\left( \left(\frac{\sqrt{2\alpha}\|x-x_0\|}{\sigma}\right)^{-(d+1)/2} \right) 
    \right\}
    . \label{eqn:occupation_asymptotics}
\end{align}
Since we only have asymptotics for the equilibrium frequency, 
we will only use the latter (asymptotic) form.

%%%%%%% %%%%%%%%%%%
\subsubsection*{Outflux of families}

Now, we can obtain the first term in equation \eqref{eqn:gestalt_q}, the (outflux of families).
Define $\gamma(r_0)$ to be this outflux; 
i.e.\ the mean number of individuals per unit time that are born carrying the $B$ allele closer than $r_0$ to the original patch
but migrate to a larger distance.
Then for any point $x$ further than $r_0$ from the patch, 
the mean occupation density $q(x)$ is equal to the integral of $\gamma$ over all locations $x_0$ at distance $r_0$ from the patch,
multiplied by the mean occupation density at $x$ for migrant families beginning at at location.
This is a more precise statement of the decomposition of the equilibrium frequency from equation \eqref{eqn:gestalt_q}:
\begin{align}
    q(x) &= \gamma(r_0) \int_{B(r_0)} u(y,x) dy,
\end{align}
so that $\gamma(r_0) = q(x) / \int_{B(r_0)} u(y,x) dy$, for any choice of $x$.
Now, using equations \eqref{eqn:eqfreq} and \eqref{eqn:occupation_asymptotics}, 
we can obtain that (see Appendix \ref{apx:lesbegue})
the outflux of families 
\begin{align}
    \gamma(r_0) \simeq 
\end{align}

Denote by $Z_t$ the number of individuals such a family has after time $t$,
and the locations of these by $X_1(t), \ldots, X_{Z_t}(t)$.
Then this mean occupation density, at $x$, is the expected number of individuals ever present at $x$. 
Since the expected number of individuals present at time $t$ is $e^{\alpha t}$
and the location of each of these individuals is marginally distributed as a  Gaussian with variance $\sigma^2 t$,
we can write this expectation as
\begin{align}
    u(x) &= \int_0^t \E[ \#\{ k : \; X_k(t) \in dx \} ] dt \\
    &= \int_0^t e^{\alpha t} \frac{1}{\sqrt{2 \pi t \sigma^2}} e^{-\frac{x^2}{2t\sigma^2}} dt \\
    &= \frac{ 1 }{ \sigma \sqrt{2 \alpha} } \exp\left( - \frac{ x \sqrt{2\alpha} }{ \sigma } \right) 
\end{align}
\plr{check that}
The integral is evaluated in Appendix~\ref{apx:some_integrals}.
\plr{Also examine non-dependence on $r_0$.}

Combining this with the solution for $q(r)$ (equation~\eqref{eqn:eqfreq}) as in \eqref{eqn:q_gestalt},
we then have that (with $C=1/2$)
\begin{align} \label{eqn:outflux}
    \gamma = \text{( outflux of families )} \approx (1/2) \rho \sigma \sqrt{2\alpha} .
\end{align}

\subsubsection*{Probability of establishment}

Suppose that the new patch is at distance $r$.
The probability density for the first time such a Brownian motion $B_s$ hits $r$ is
\citep[XXX]{feller}
\plr{work in $d=2$}
% note $\sigma B_t \deq B_{\sigma^2 t}$ so $\P\{ \sup_{s\le t} \sigma B_s \le r \} = \P\{ \sup_{s \le \sigma^2 t} B_s \le r \}
\begin{align} \label{eqn:hitting_dist}
  \P\{ B_t=r \;\mbox{and} \; B_u<r \;\mbox{for all}\; u<t\} =  \frac{r}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{r^2}{2t\sigma^2}\right) .
\end{align}
Let $\tau$ be this random time at which a given migrant family hits the new patch,
with the convention that $\tau =\infty$ if the family never gets here.
We approximate the chance that a given family arrives at the new patch at time $t$, i.e.\ that $\tau=t$,
by the product of equation~\eqref{eqn:hitting_dist} and the probability that the migrant family survives until time $t$.
For convenience, define $\alpha := - \log(1-s_m)$ to be the rate of decrease between patches (note that $\alpha \approx s_m$),
and $\mu_K = \E[K]$ to be the limiting mean family size conditioned on survival,
so that $1-p_e(t) \simeq (1/\mu_K) \exp(-\alpha t)$.
Then, the probability distribution of $\tau$ is given by
\begin{align}
    \P\{ \tau \in dt \} =  \frac{1}{\mu_K} \frac{r}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{r^2}{2t\sigma^2} - \alpha t \right) dt  .
\end{align}

By integrating over~$t$, the chance that this family ever arrives at the new patch is
\begin{align} 
\label{eqn:estab_prob}
  \begin{split}
      \P\{ \tau < \infty \} &= \int_0^\infty \P\{ \tau \in dt \} = \frac{1}{\mu_K} \exp\left( - \frac{ r \sqrt{2 \alpha}}{\sigma} \right) .
  \end{split}
\end{align}
This is, up to the constant $\mu_K$, the chance that a Brownian motion killed at rate $\alpha$
ever hits $r$ \citep{feller}.
This approximation depends on $r$ being large enough so that the approximation $1-p_e(t) \approx (1-s_m)^t/\mu_K$ holds.

Now, each member of a family that successfully arrives at the new patch
(where the $B$ allele is again beneficial)
has a chance $p_e$ of establishing.
The precise distribution of the number of members of a family that enter the patch
given the trunk arrives there
depends on the geometry of the patch,
but since family members should not be too far from the trunk,
we can approximate this by the total size of the family.
If the family has $K$ individuals when the trunk arrives at the new patch,
the chance that any establish is approximately $1-(1-p_e)^K$;
and so that chance that the family establishes, given it arrives at the new patch, 
is 
\begin{align}
    \label{eqn:prob_estab_arrival}
    \E[1-(1-p_e)^K] \approx 2 s_p \mu_K/\xi^2 .
\end{align}

The probability of establishment of an outmigrant from the already-established patch is 
equal to the probably the trunk arrives (equation~\eqref{eqn:estab_prob}) multiplied by the probability it establishes (equation~\eqref{eqn:prob_estab_arrival}),
namely
\begin{align} \label{eqn:bp_prob_estab}
    F(r) &:= \text{( prob family establishes at $r$ )} = \frac{2s_p}{\xi^2} \exp\left( - \frac{\sqrt{2 \alpha} r}{\sigma} \right) ,
\end{align}

\subsubsection*{Putting it all together}

Combining this with \eqref{eqn:bp_prob_estab}, as in \eqref{eqn:gestalt_q}, 
the rate of establishment, for a patch of area $A$ at distance $r$ from an already occupied patch, is approximately
\begin{align} % \label{eqn:migrate}
    \migrate(r) = \gamma F(r) = \frac{\rho \sigma s_p\sqrt{2s_m} }{\xi^2}  \exp\left( -\frac{ \sqrt{2 s_m} r}{\sigma} \right).
\end{align}
Note that the exponential term is expected based on \eqref{eqn:eqfreq} -- 
since establishment requires rare long distant migrants to be present, and \eqref{eqn:eqfreq} gives the density of such --
but the first term required other arguments.

If this establishment rate is low, 
the time (in generations) before migrant copies of the allele successfully establish through migration in the second patch
is approximately exponentially distributed with mean $1/\migrate$.
Neglecting the time for such a migrant allele to establish once present in the second patch,
the time scale on which migrant alleles transit and fix between patchs is $1/\migrate$.

@@

As noted above, the length of the hitchhiking segment depends on the time
spent ``in transit'' between the patches.
Above, we approximated this by the time $\tau$ it takes the trunk lineage to hit the new patch,
except that studying a successful migrant family means we need to condition on the establishment occurring, i.e.\ on $\tau < \infty$.
By the above arguments, and equation~\eqref{eqn:estab_prob}, we have that the conditional probability density is
\begin{align}
  \P\{\tau\in dt | \tau<\infty \} &= \frac{\P\{\tau \in dt\}}{\P\{\tau<\infty\}} \\
  &= \frac{r}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{r^2}{2t\sigma^2} -\alpha t + \sqrt{\alpha} \frac{r}{\sigma} \right) .
\end{align}
\plr{more dimensions}
Since each generation spent in transit provides an opportunity for recombination,
if recombination is Poisson, the length of the hitchhiking segment (in Morgans) on each side of the site will be exponentially distributed
with mean $\tau$, so that if $L$ is the length on, say, the right, then
\begin{align} \label{eqn:haplen_cdf}
\P\{L>\ell\} &= \E[e^{-\ell \tau}|\tau<\infty] \\
  &= \exp\left\{{-\frac{r}{\sigma}\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} .
\end{align}
This is also the Laplace transform of $\tau$;
we can differentiate it to find that $\E[\tau] = r/(\sigma\sqrt{2\alpha})$
and $\var[\tau] = r/( 2^{3/2}\alpha^{3/2}\sigma)$.
Furthermore, this implies that if $Y$ is an exponential random variable with rate $r\sqrt{2}/\sigma$,
then $L$ has the same distribution as $(Y + \sqrt{\alpha})^2 - \alpha$,
from which we can obtain that
$\E[L] = \sigma \sqrt{2\alpha}/r + \sigma^2/r^2$ and $\var[L] = 2\alpha\sigma^2/r^2 + 4 \sigma^3 \sqrt{2\alpha}/r^3 + 5 \sigma^4 / r^4$.

@@

In the above we already assumed that the inter-patch distance $r$ is large relative to the migration-selection cline width $\sigma/\sqrt{s_m}$;
by taking the large-$r$ limit above we can obtain a more intuitive description of the process
via a central limit for $\tau$.
Define $\eta_r$ to be the deviation of $\tau$ from its mean, normalized by its standard deviation,
so that $\tau = r/\sigma(2\alpha)^{-1/2} + \eta_r \sqrt{r/\sigma} (2\alpha)^{-3/4}$.
Plugging this into equation~\eqref{eqn:haplen_cdf},
after some algebra (and using that $\sqrt{r+\epsilon} = \sqrt{r} + \frac{\epsilon}{2\sqrt{r}} - \frac{\epsilon^2}{2\sqrt{r}^3} + O(\epsilon^3)$) we obtain that
\begin{align}
  \E[e^{-\ell \eta}] &= \exp\left\{ \frac{\ell^2}{2} + O\left(\frac{1}{r}\right) \right\}.
\end{align}
By the properties of Laplace transforms \citep{durrett},
this implies that $\eta_r$ converges in distribution to a standard Gaussian random variable as $r \to \infty$.
In other words, we now know that the distribution of $\tau$ about its mean is approximately Gaussian.
Since the standard deviation of $\tau$ is of lower order than its mean,
a similar calculation shows that for large $r$, the haplotype halflength $L$ is approximately exponentially distributed, 
with mean $\sqrt{2\alpha} \sigma/r$.

% \begin{align}
%   \E[e^{-\ell \eta}] &= \E\left[ \exp\left\{-\ell\left(\tau - \frac{r}{\sigma\sqrt{2\alpha}}\right)\frac{(2\alpha)^{3/4}\sqrt{\sigma}}{\sqrt{r}} \right\} \right] \\
%   &= \E\left[ \exp\left\{ - \left(\frac{\ell (2\alpha)^{3/4} \sqrt{\sigma}}{\sqrt{r}}\right) \tau \right\} \right] \exp\left( (2\alpha)^{1/4} \sqrt{\frac{r}{\sigma}} \right) \\
% \end{align}

@@

@@

>>>>

Let $p(x)$ be the probability that a new mutant $B$ allele arising at location $x$ 
relative to the patch fixes within the poor habitat patch.
Under various assumptions, precise expressions can be found for $p(x)$ \citep{barton1987establishment},
% the function $p(x)$ can be found (in one dimension) in terms of Jacobi elliptic functions,
but these are not particularly helpful.
The total successful mutational influx per generation is given by the integral of $p(x)$ over the entire species range,
% $\int \rho \mu p(x) dx$,
and hence depends in a complicated way on the relationship of patch width and selection coefficients,
but still scales linearly with the mutational influx density $\rho \mu$.
If the width of the patch is large relative to $\sigma/\sqrt{2s_m}$, 
then a reasonable approximation is to assume that $p(x) = p_e$ within the patch, and $p(x) = 0$ otherwise,
for a patch of area $A$ approximating $\int p(x) dx \approx 2 s_p A / \xi^2$. 
This holds for large patches, in which the rate at which mutations arise within the patch is larger 
than the rate at which they appear nearby;
we examine this more generally via simulation in Appendix \ref{apx:establishment_sims}.

@@

Migration--selection balance ensures that 
there will be some number of $B$ alleles present in the regions where they are disadvantageous,
but only very few, and rarely, far away from the patch where $B$ is established.
Denote the expected frequency of allele $B$ at displacement $x$ relative to the patch by $q(x)$.
Following \citet{slatkin1973geneflow}, one can write down a differential equation to which $q(x)$ is the solution,
which in one dimension can be solved to give $q(x)$ in terms of Jacobi elliptic functions (see Appendix \ref{apx:elliptic_integrals}).
More general, asymptotics of the solution implies that for large $|x|$ the mean occupation frequency decays exponentially, 
with a square-root correction in two dimensions (see Appendix \ref{apx:asymptotics}): 
\begin{align} \label{eqn:eqfreq}
  q(x) \approx C \left( |x| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |x| \sqrt{2 s_m} / \sigma) \qquad \text{for large $|x|$},
\end{align}
where $d$ is the dimension ($d=1$ or $2$), and $C$ is a constant depending on the geographic shape of the populations and the selection coefficients.
In applications we fit $C$ to the data;
to get concrete numbers from theory we take $C=1$ if necessary.
From simulations, $C$ seems to be not too far from 1 in practice (see Appendix \ref{apx:patchy_sims}).

This expected frequency $q(r)$ is the \emph{time-averaged} occupation frequency,
or in other words, the total number of $B$ alleles found at distance $r$ across $T$ generations, divided by $T$.
Therefore, if we look at a small area of size $dx$, 
then $q(r) dx$ is equal to the probability of finding any $B$ alleles there,
multiplied by the mean number of such, if there are any.
This points to an important fact: 
close to the original patch, the ``equilibrium frequency'' $q(r)$ describes well the density of the $B$ allele at most times,
but far away from the patch, 
the equilibrium is composed of rare excursions of a family of $B$ alleles, interspersed by long periods of absence.
The relevant time scale on which $B$ alleles migrate between patches is given by the rate of appearance of such families.
To compute the rate of establishment by such migration,
we use properties of subcritical branching processes and Brownian motion,
combined with equation \eqref{eqn:eqfreq},
to obtain an approximation that is valid if the distance $r$ is large (at least five times $\sigma/\sqrt{s_m}$, say).

@@

Now we turn to the expression \eqref{eqn:gestalt_migrate} for the outflux of successful migrants.
Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch,
which we assume is circular, has relatively small area $A$ and is located near position $y$.
% The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-p_e(t))$,
% multiplied by the probability that the family reaches the patch, given survival.
% If $t$ is large then as in the previous section, $(1-p_e(t)) \approx e^{-\alpha t}/\E[K]$.
It would be possible to analyze directly the probability that a given migrant family establishes in $A$,
as in \citet{barton1987establishment,mckean1975KPP};
but we make a simpler approximation, approximating this by the chance that the ``trunk'' hits the new patch,
multiplied by the chance that at least member of the family escapes demographic stochasticity to establish.
Write $h(x,y)$ for the probability that the Brownian trunk
hits the new patch centered at $y$ before the family dies out,
and $f(A)$ for the chance that the family manages to establish in the new patch,
given that it successfully arrives.
As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is an integral of $h(x,y)$:
\begin{align}\label{eqn:migrate_integral}
  \migrate(y) &= \rho \int_{\partial B(r_0)} \gamma(x) h(x,y) f(A) dx .
\end{align}

Intuitively, $h(x,y)$ counts each family once if it hits the patch,
while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
Since the expected total time spent in the patch does not depend on the distance at which the family started
if we know that it survived to hit the patch,
it is intuitively reasonable that $u(x,y)$ and $h(x,y)$ are simply related.
Using this fact,
and that $(1-p_e(t)) \approx e^{-\alpha t}/\E[K]$,
it turns out (see Appendix~\ref{apx:hitting_occupation}) that
\begin{align} 
  h(x,y) = \frac{ s(A) }{ \E[K] } u(x,y),
\end{align}
where $s(A)$ is the total expected occupation time of the patch given the family arrives,
and recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a constant factor
(at least asymptotically for large distances),
we now know that the outflux of successful migrant families is equal to the mean occupation frequency
multiplied by $s(A) f(A) / \E[K]$.
The probability $f(A)$ that the family does not die out once it arrives at the new patch
is if each member arriving has chance $p_e$ of establishing, independently, $\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$.
% If we let $w$ be the width of the patch, then the remaining term is
% \begin{align}
%   s(w) \simeq \frac{1}{2\alpha} \left( 1 - \left( \frac{\pi w \sqrt{2 \alpha}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2\alpha}/\sigma} \right).
% \end{align}
Combining these terms with the expression for $s(A)$ from Appendix~\ref{apx:hitting_occupation},
\begin{align}
  \label{eqn:migrate}
  \migrate(y) \approx q(y) \times \rho \frac{ s_p  }{ \alpha \xi^2 } 
  % \left( 1 - \left( \frac{\pi w \sqrt{2 \alpha}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2\alpha}/\sigma} \right).
\end{align}
For patches that are small relative to $\sigma/\sqrt{2s_m}$, there is an additional correction term (see Appendix~\ref{apx:hitting_occupation}),
but small patches violate other assumptions above, so we omit it.

@@

As noted above, the length of the hitchhiking segment depends on the time
spent ``in transit'' between the patches.
Above we argued that a good model for this transit is a single Brownian trunk lineage
surrounded by a cloud of close relatives of random size $K$
whose chance of surviving until $t$ generations is $p_e(t) \simeq e^{-\alpha t} / \E[K]$.
Consider a single such family, and let $\tau$ be the (random) time it first hits the new patch,
with $\tau = \infty$ if it never does.
By pulling out the constant $1/\E[K]$, we can use standard results on hitting times of $d$-dimensional Brownian motion
that is killed at rate $\alpha$.
In particular, we know that if the patch is circular with radius $w$, then
using \citet[][(2.2.0.1 and 4.2.0.1)]{borodin2002handbook}
\begin{align}
  \E[e^{-\ell \tau}] &=
    \begin{cases}
      e^{- R \sqrt{2(\alpha+\ell)}/\sigma} \qquad & d=1 \\
      \frac{ K_0( (R+w)\sqrt{2(\alpha+\ell)}/\sigma) }{ K_0( w\sqrt{2(\alpha+\ell)}/\sigma) } \qquad & d=2 
    \end{cases} 
    % &= \left( R/w +1 \right)^{-\frac{d-1}{2}} e^{- R \sqrt{2(\alpha+\ell)}/\sigma}  (1+O(1/w)) .
\end{align}
In particular, we can find 
$\P\{\tau < \infty\} = \lim_{\ell \to 0} \E[\exp(-\ell \tau)]$. % \simeq (R/w+1)^{-(d-1)/2} \exp(- R \sqrt{2\alpha}/\sigma)$.

% Above, we approximated this by the time $\tau$ it takes the trunk lineage to hit the new patch;
% now that we are studying a successful migrant family means we need to condition on the establishment occurring, i.e.\ on $\tau < \infty$.
% By the above arguments, and equation~\eqref{eqn:estab_prob}, we have that the conditional probability density is
% \begin{align}
%   \P\{\tau\in dt | \tau<\infty \} &= \frac{\P\{\tau \in dt\}}{\P\{\tau<\infty\}} \\
%   &= \frac{R}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{R^2}{2t\sigma^2} -\alpha t + \sqrt{\alpha} \frac{R}{\sigma} \right) .
% \end{align}
% \plr{more dimensions}

Since each generation spent in transit provides an opportunity for recombination,
if recombination is Poisson, the length of the hitchhiking segment (in Morgans) on each side of the site will be exponentially distributed
with mean $\tau$, so that if $L$ is the length on, say, the right, then in one dimension
\begin{align} \label{eqn:haplen_cdf}
\P\{L>\ell\} &= \E[e^{-\ell \tau}|\tau<\infty] \\
  &= \exp\left\{{-\frac{R}{\sigma}\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} .
\end{align}
This is also the Laplace transform of $\tau$;
we can differentiate it to find that $\E[\tau] = (R\sqrt{2\alpha}/\sigma)\times 1/(2\alpha)$
and $\var[\tau] = (R\sqrt{2\alpha}/\sigma) \times 1/(2\alpha)^2$.
Furthermore, this implies that if $Y$ is an exponential random variable with rate $R\sqrt{2}/\sigma$,
then $L$ has the same distribution as $(Y + \sqrt{\alpha})^2 - \alpha$,
from which we can obtain that
$\E[L] = \sigma \sqrt{2\alpha}/R + \sigma^2/R^2$ and $\var[L] = 2\alpha\sigma^2/R^2 + 4 \sigma^3 \sqrt{2\alpha}/R^3 + 5 \sigma^4 / R^4$.
For two dimensions, asymptotics of Bessel functions show that $L$ has the same tail behavior,
but how other properties might change is unclear.

@@

In the above we already assumed that the inter-patch distance $R$ is large relative to the migration-selection cline width $\sigma/\sqrt{s_m}$;
by taking the large-$R$ limit above we can obtain a more intuitive description of the process
via a central limit for $\tau$.
Define $\eta_\tau$ to be the deviation of $\tau$ from its mean, normalized by its standard deviation,
so that $\tau = R/\sigma(2\alpha)^{-1/2} + \eta_\tau \sqrt{R/\sigma} (2\alpha)^{-3/4}$.
Plugging this into equation~\eqref{eqn:haplen_cdf},
after some algebra (and using that $\sqrt{R+\epsilon} = \sqrt{R} + \frac{\epsilon}{2\sqrt{R}} - \frac{\epsilon^2}{2\sqrt{R}^3} + O(\epsilon^3)$) we obtain that
\begin{align}
  \E[e^{-\ell \eta}] &= \exp\left\{ \frac{\ell^2}{2} + O\left(\frac{1}{R}\right) \right\}.
\end{align}
Since the Laplace transform uniquely determines a distribution \citep{durrett1996probability},
this implies that $\eta_\tau$ converges in distribution to a standard Gaussian random variable as $R \to \infty$.
In other words, we now know that the distribution of $\tau$ about its mean is approximately Gaussian.
Since the standard deviation of $\tau$ is of lower order than its mean,
a similar calculation shows that for large $R$, the haplotype halflength $L$ is approximately exponentially distributed, 
with mean $\sqrt{2\alpha} \sigma/R$.

% \begin{align}
%   \E[e^{-\ell \eta}] &= \E\left[ \exp\left\{-\ell\left(\tau - \frac{R}{\sigma\sqrt{2\alpha}}\right)\frac{(2\alpha)^{3/4}\sqrt{\sigma}}{\sqrt{R}} \right\} \right] \\
%   &= \E\left[ \exp\left\{ - \left(\frac{\ell (2\alpha)^{3/4} \sqrt{\sigma}}{\sqrt{R}}\right) \tau \right\} \right] \exp\left( (2\alpha)^{1/4} \sqrt{\frac{R}{\sigma}} \right) \\
% \end{align}

XXX Check what happens in two dimensions. XXX

@@

If $q(r)$ is a radially symmetric equilibrium frequency and $\xi(t,x)=q(|x|)$,
then $\partial_t \xi(t,x) = 0$.
So, using \eqref{eqn:radial_fkpp}, a radially symmetric solution 
with $\gamma(r) = -s < 0$ for all $r>r_0$ 
solves, for $r_0 < r < \infty$,
\begin{align}
    \partial_r^2 q(r) + \frac{d-1}{r} \partial_r q(r) - \frac{2s}{\sigma^2} q(r) (1-q(r)) = 0 \\
    \lim_{r \to \infty} q(r) = \lim_{r \to \infty} \partial_r q(r) = 0 \qquad 
    0 < q(r) < 1 
\end{align}
Since $q(r) \to 0$ as $r \to \infty$, so $q(r) (1-q(r)) \approx q(r)$,
it can be shown 
% theory in \citet{wasow1965asymptotic} implies 
that the true equilibrium frequency $q$ is close, for large $r$, to the solution to
\begin{align}
    \partial_r^2 u(r) + \frac{d-1}{r} \partial_r u(r) - \frac{2s}{\sigma^2} u(r) = 0  \label{eqn:bessel} \\
    \lim_{r \to \infty} u(r) = \lim_{r \to \infty} \partial_r u(r) = 0 \qquad  .
\end{align}
This has general solution given by a modified Bessel function:
using \citet{gradshteyn2007table} 8.494.9,
the general solution is
\begin{align}
    u(r) = C' (r-r_0)^{(2-d)/2} K_{(2-d)/2} \left( (r-r_0) \sqrt{2s}/\sigma \right) ,
\end{align}
where $C'$ and $r_0$ are chosen to match boundary conditions.
Asymptotics of Bessel functions \citep[][8.451.6]{gradshteyn2007table} then imply that
\begin{align}
    q(r) \approx C r^{(1-d)/2} \exp \left( -r \sqrt{2s}/\sigma \right) + O(1/r),
\end{align}
where $C$ is a different constant.

\subsection{Doing the integrals}
\label{apx:some_integrals}

@@

\section{Temporary work: hitting probs}
